---
title: DS202 Lab2 Team3
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(classdata)
```

# Lab report #1

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

## step 1 (inspect the first few lines of the data set: what variables are there? of what type are the variables? what does each variable mean? what do we expect their data range to be?)

```{r}
# loading data
data(ames)

# insepcting data
head(ames)
names(ames)
```
As a team, we found that there are 16 variables in total. “Parcel ID”, “Address”, “Multi Sale”, “AC”,  and “FirePlace” are chr variables. “Style”, “Occupancy”, and “Neighborhood” are fctr variables. “Sale Price”, YearBuilt”, “Acres”, “TotalLivingArea (sf)”, “Bedrooms”, “FinishedBsmtArea (sf)”, and LotArea(sf) are dbl variables. Lastly, “Sale Date” is a date variable. Most of these variables are self-explanatory based on their names. 
## step 2 (is there a variable of special interest or focus? Let's call it the "main variable")

<<<<<<< HEAD
=======
```{r}
# Finding Sales Price
summary(ames$`Sale Price`)

# Range of Ames Sales Price
range(ames$`Sale Price`, na.rm = TRUE)

```
As a group, we found that the main variable of interest is sales price, which represents the final selling price of houses in Ames. This variable helps us understand how different housing features like size, bedrooms, neighborhood, and year built may affect the price of a home.

>>>>>>> 5c7ff84bd8d844d19a7ad284be6c890a4f8e0314
## step 3 (start the exploration with the main variable: what is the range of this variable? draw a histogram for a numeric variable or a bar chart, if the variable is categorical. what is the general pattern? is there anything odd?)

## step 4 (pick a variable that might be related to the main variable. what is the range of that variable? plot. describe the pattern. what is the relationship to the main variable? plot a scatterplot, boxplot or facetted barcharts dependening on the types of variables involved. Describe overall pattern, does this variable describe any oddities discovered in 3? Identify/follow-up on any oddities.)

Tyler's work:
```{r}
#my variable is TotalLivingArea (sf)

range(ames$`TotalLivingArea (sf)`, na.rm = TRUE)
# ^this^ function shows the range is 0-6007

#bar graph with the total living area binned
library(dplyr) 
library(ggplot2) 
ames %>% 
  mutate( LivingAreaBin = cut(`TotalLivingArea (sf)`, breaks = 6, dig.lab = 6) 
          ) %>% 
  group_by(LivingAreaBin) %>% 
  summarise(AvgSalePrice = mean(`Sale Price`, na.rm = TRUE)) %>% 
  ggplot(aes(x = LivingAreaBin, y = AvgSalePrice)) + geom_col(fill = "steelblue") + labs( title = "Average Sale Price by Living Area", x = "Total Living Area (binned)", y = "Average Sale Price" ) + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
<br> From this bar plot, we can see that the lower total living area houses tend to have a higher average sale price. I find it odd that the higher living area houses are cheaper, as you would think that the bigger houses would be more expensive. 

Jacob's work:

Favour's work:

Kavya's work:
